---
title: "Testing atlantisom: generate survey sampled length comps"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

## Introduction

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. 

The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models can be run using different climate forcing, fishing, and other scenarios. Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario. Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.

Initial testing was conducted by S. Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R". Initial tests are expanded and documented in more detail in these pages. C. Stawitz improved and streamlined the setup and intialization sections. 

## Setup

First, you will want to set up libraries and install atlantisom if you haven't already. This document assumes atlantisom is already installed. For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html). 

This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes]. 

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```


## Initialize input files and directories, read in "truth"

Abbreviated here; for a full explanation please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html). This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

if(initCCA) source(here("config/CCConfig.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

```


```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```


```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  truth <- result
} 

if(initNEUS) {
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

if(initNOBA){
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

```

## Simulate a survey part 4: sample for length composition

This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:

 *  age samples from `calc_stage2age` and `sample_ages`
 *  length samples from `calc_age2length` and `sample_lengths` (TO BE WRITTEN)
 *  weight samples from `calc_age2length` and`sample_weights` (TO BE WRITTEN)
 *  diet samples from `sample_diet`
 
Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly. The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition. The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function. We determined that it will not work better to create a length comp for the whole population, which could then be sampled. We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document. 

### Standard survey settings for testing 
To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species. 

```{r sppgroups}

# make defaults that return a standard survey, implement in standard_survey
# users need to map their species groups into these general ones
#   large pelagics/reef associated/burrowers/otherwise non-trawlable
#   pelagics
#   demersals
#   selected flatfish
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

if(initCCA) { #Sarah's CCA Grouping
  nontrawl <- c("Shark_C","Yelloweye_rockfish","Benthopel_Fish","Pisciv_S_Fish",
                "Pisciv_T_Fish","Shark_D","Shark_P")
  pelagics <- c("Pisciv_V_Fish","Demersal_S_Fish","Pacific_Ocean_Perch","Mesopel_M_Fish",
                "Planktiv_L_Fish","Jack_mackerel","Planktiv_S_Fish","Pacific_sardine",
                "Anchovy","Herring","Pisciv_B_Fish")
  demersals <- c("Demersal_P_Fish","Planktiv_O_Fish","Demersal_D_Fish",
                 "Demersal_DC_Fish","Demersal_O_Fish","Darkblotched_rockfish",
                 "Demersal_F_Fish","Demersal_E_Fish","Bocaccio_rockfish",
                 "Demersal_B_Fish","Shark_R","Mesopel_N_Fish","Shark_B","Spiny_dogfish",
                 "SkateRay")
  selflats <- c("Pisciv_D_Fish", "Arrowtooth_flounder","Petrale_sole")
}

if(initNEUS) { # Sarah's NEUS Grouping
  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", 
                 "Demersal_S_Fish","Demersal_B_Fish","Demersal_DC_Fish",
                 "Demersal_O_Fish","Demersal_F_Fish",
                 "Shark_B", "SkateRay")
  selflats <- c("Pisciv_B_Fish")
}

if(initNOBA) { # Sarah's NOBA Grouping
  nontrawl <- c("Sharks_other", "Pelagic_large","Mesop_fish")
  pelagics <- c("Pelagic_small","Redfish_other","Mackerel","Haddock",
                "Saithe","Redfish","Blue_whiting","Norwegian_ssh","Capelin")
  demersals <- c("Demersals_other","Demersal_large","Flatfish_other","Skates_rays",
                 "Green_halibut","North_atl_cod","Polar_cod","Snow_crab")
  selflats <- c("Long_rough_dab")
}

```

The following settings are for our example standard survey once per year, most areas, with mixed efficiency and selectivity: 

```{r stdbtsurvey-spec, message=FALSE, warning=FALSE}

# general specifications for bottom trawl survey, with items defined above commented out to avoid wasting time loading already loaded files:
#   once per year at mid year
# generalized timesteps all models
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

midptyr <- round(median(seq(1,stepperyr)))

annualmidyear <- seq(midptyr, noutsteps, stepperyr)

#   ~75-80% of boxes (leave off deeper boxes?)
boxpars <- load_box(d.name, box.file)
boxsurv <- c(2:round(0.8*(boxpars$nbox - 1)))

#   define bottom trawl mixed efficiency
ef.nt <- 0.01 # for large pelagics, reef dwellers, others not in trawlable habitat
ef.pl <- 0.1  # for pelagics
ef.dm <- 0.7  # for demersals
ef.fl <- 1.1  # for selected flatfish

# bottom trawl survey efficiency specification by species group
effnontrawl <- data.frame(species=nontrawl, efficiency=rep(ef.nt,length(nontrawl)))
effpelagics <- data.frame(species=pelagics, efficiency=rep(ef.pl,length(pelagics)))
effdemersals <- data.frame(species=demersals, efficiency=rep(ef.dm,length(demersals)))
effselflats <- data.frame(species=selflats, efficiency=rep(ef.fl,length(selflats)))

efficmix <- bind_rows(effnontrawl, effpelagics, effdemersals, effselflats)

#   mixed selectivity (using 10 agecl for all species)
#     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
#     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
#     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10

sigmoid <- function(a,b,x) {
  1 / (1 + exp(-a-b*x))
}

# survey selectivity specification by species group
selnontrawl <- data.frame(species=rep(nontrawl, each=10),
                          agecl=rep(c(1:10),length(nontrawl)),
                          selex=rep(1.0,length(nontrawl)*10))
selpelagics <- data.frame(species=rep(pelagics, each=10),
                          agecl=rep(c(1:10),length(pelagics)),
                          selex=sigmoid(5,1,seq(-10,10,length.out=10)))
seldemersals <- data.frame(species=rep(demersals, each=10),
                          agecl=rep(c(1:10),length(demersals)),
                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))
selselflats <- data.frame(species=rep(selflats, each=10),
                          agecl=rep(c(1:10),length(selflats)),
                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))

selexmix <- bind_rows(selnontrawl, selpelagics, seldemersals, selselflats)

# use this constant 0 cv for testing
surv_cv_0 <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

#   define bottom trawl survey cv by group
cv.nt <- 1.0 # for large pelagics, reef dwellers, others not in trawlable habitat
cv.pl <- 0.5  # for pelagics
cv.dm <- 0.3  # for demersals
cv.fl <- 0.3  # for selected flatfish

# specify cv by species groups
surv_cv_nontrawl <- data.frame(species=nontrawl, cv=rep(cv.nt,length(nontrawl)))
surv_cv_pelagics <- data.frame(species=pelagics, cv=rep(cv.pl,length(pelagics)))
surv_cv_demersals <- data.frame(species=demersals, cv=rep(cv.dm,length(demersals)))
surv_cv_selflats <- data.frame(species=selflats, cv=rep(cv.fl,length(selflats)))

surv_cv_mix <- bind_rows(surv_cv_nontrawl, surv_cv_pelagics, surv_cv_demersals, surv_cv_selflats)

```

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition. 

```{r stdsurveyNbased}

survey_testNstd <- create_survey(dat = truth$nums,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv,
                                 effic = efficmix,
                                 selex = selexmix)

# consider saving this interim step if it takes a long time go generate


```

Next we apply the aggregateDensityData function can to resn and structn for survey times, species, and boxes.

```{r aggdens}

# aggregate true resn per survey design
survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv)

# aggregate true structn per survey design
survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxsurv)

```

Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey. The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call. 

```{r stdsurvey-lensamp, warning=FALSE, message=FALSE}

#   define n fish for biological sampling by group
#   this could easily be a vector or time series, constant here
ns.nt <- 25 # for large pelagics, reef dwellers, others not in trawlable habitat
ns.pl <- 1000  # for pelagics
ns.dm <- 1000  # for demersals
ns.fl <- 1000  # for selected flatfish

effNnontrawl <- data.frame(species=nontrawl, effN=rep(ns.nt,length(nontrawl)))
effNpelagics <- data.frame(species=pelagics, effN=rep(ns.pl,length(pelagics)))
effNdemersals <- data.frame(species=demersals, effN=rep(ns.dm,length(demersals)))
effNselflats <- data.frame(species=selflats, effN=rep(ns.fl,length(selflats)))

effNmix <- bind_rows(effNnontrawl, effNpelagics, effNdemersals, effNselflats)

# apply default sample fish as before to get numbers
numsstd <- sample_fish(survey_testNstd, effNmix)

#dont sample these, just aggregate them using median (effNmix does nothing)
structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)

resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)
```

```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE}

# now cut these down to a single species for testing

atf_numsstd <- numsstd[numsstd$species == "Arrowtooth_flounder",]
atf_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
atf_resnstd <- resnall[resnstd$species == "Arrowtooth_flounder",]

atf_length_stdsurv <- calc_age2length(structn = atf_structnstd,
                                 resn = atf_resnstd,
                                 nums = atf_numsstd,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)


```

Plot samples one species:

```{r atflengthsamp1-test}

lfplot <- ggplot(atf_length_stdsurv$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")

# try dir = "v" option for vertical lf comparisons

```

## Samples for all species, then SAVE

Now run for all species in the survey.

```{r allspplengthsamp}

length_stdsurv <- calc_age2length(structn = structnstd,
                                 resn = resnstd,
                                 nums = numsstd,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

#save for later use, takes a long time to generate
saveRDS(length_stdsurv, file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))

```

Runtime for census lengths was 5.5 hours for CCA. For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.

Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html): 

```{r loadlengthcomp, echo=TRUE, eval=FALSE}

length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

length_stdsurv <- readRDS(file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))

```

Comparison plots for other species:

```{r moreplots, echo=TRUE, eval=FALSE}

ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
  filter(species == "Herring")

ss_length_stdsurv <- length_stdsurv$natlength %>%
  filter(species == "Herring")

# add sample, this is census, probably want it first?
lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")

```


## References
