---
title: "Testing atlantisom: true biomass output from California Current Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')

library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)

```

## Introduction

This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

if(initCCA){
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  functional.groups.file <- "CalCurrentV3Groups.csv"
  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
  biol.prm.file <- "CalCurrentV3_Biol.prm"
  box.file <- "CalCurrentV3_utm.bgm"
  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
  run.prm.file <- "CalCurrentV3_run.xml"
  scenario.name <- "CCV3"
}

if(initNEUS){
  d.name <- here("atlantisoutput","NEUStest20160303")
  functional.groups.file <- "NeusGroups.csv" 
  biomass.pools.file <- ""
  biol.prm.file <- "at_biol_neus_v15_DE.prm"
  box.file <- "neus30_2006.bgm"
  initial.conditions.file <- "inneus_2012.nc"
  run.prm.file <- "at_run_neus_v15_DE.xml"
  scenario.name <- "neusDynEffort_Test1_"
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  functional.groups.file <- "nordic_groups_v04.csv" 
  biomass.pools.file <- "nordic_biol_v23.nc"
  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
  box.file <- "Nordic02.bgm"
  initial.conditions.file <- "nordic_biol_v23.nc"
  run.prm.file <- "nordic_run_v01.xml"
  scenario.name <- "nordic_runresults_01"
}
# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this. Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case. For now, NOBA filenames have been changed to include prefix "output"
```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  CCAresults <- result
} 

if(initNEUS) {
  d.name <- here("atlantisoutput","NEUStest20160303")
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  NEUSresults <- result
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  NOBAresults <- result
}

```


## Simulate a survey part 1: census to compare with truth

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html). The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages: 

```{r census-spec, echo=TRUE}

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
# BUT CHECK if newer Atlantis models can do age-specific outputs
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# these are model specific, generalized above
# if(initCCA) boxall <- c(0:88) 
# if(initNEUS) boxall <- c(0:29)
# if(initNOBA) boxall <- c(0:59) 

# should return all model output timesteps; need to generalize
if(initCCA) timeall <- c(0:100) 
if(initNEUS) timeall <- c(0:251)
if(initNOBA) timeall <- c(0:560) 
  
# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names 

# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
if(initNEUS) survspp <- funct.group.names[1:21]
if(initNOBA) survspp <- funct.group.names[1:36]
  
```

Because the results of `run_truth` provide both biomass at age and numbers at age, we can use `create_survey` on both with some modifications. The following uses the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary. Because we are making a census for testing, the survey cv argument is set to 0. 

```{r surveyBbased, eval=FALSE}

# this uses result$biomass_ages to sample biomass directly

if(initCCA) datB <- CCAresults$biomass_ages
if(initNEUS) datB <- NEUSresults$biomass_ages
if(initNOBA) datB <- NOBAresults$biomass_ages

survey_testBall <- create_survey(dat = datB,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

# call sample_survey_biomass with a bunch of 1s for weight at age
# in the code it multiplies atoutput by wtatage so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(survspp, each=10),
                    agecl=rep(c(1:10),length(survspp)),
                    wtAtAge=rep(1.0,length(survspp)*10))

surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)

#save for later use, takes a long time to generate
saveRDS(surveyB_frombio, file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

if(initCCA) CCAsurveyB_frombio <- surveyB_frombio
if(initNEUS) NEUSsurveyB_frombio <- surveyB_frombio
if(initNOBA) NOBAsurveyB_frombio <- surveyB_frombio

```

```{r readsurvey1}

if(initCCA) {
  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNEUS) {
  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNOBA) {
  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

```

Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match. Note that the our (census) survey may have more sampling in time than the Atlantis output file.

```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
# plot some comparisons with Atlantis output

# read Atlantis output files
if(initCCA) {
  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- CCAsurveyB_frombio
}

if(initNEUS) {
  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- NEUSsurveyB_frombio
}

if(initNOBA) {
  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- NOBAsurveyB_frombio
}

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- groupslookup %>%
  filter(IsTurnedOn > 0)

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% levels(surveyB_frombio$species))

#all species comparison, time intervals hardcoded for NEUS and NOBA
# WHOOPS! This is CCA, really need to generalize this but hardcoded for now
compareB <-ggplot() +
  geom_line(data=surveyB_frombio, aes(x=time,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 

```

```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 

```

```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 

```

```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 

```

```{r matchBp5, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free") 

```

After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age. This is more complex, but may be closer to the way some assessments handle survey observations. 

```{r wtagecalc, fig.cap="Check weight at age generated from biomass_age\\$atoutput divided by nums\\$atoutput for up to 64 groups.", fig.show="hold"}

if(initCCA) datN <- CCAresults$nums
if(initNEUS) datN <- NEUSresults$nums
if(initNOBA) datN <- NOBAresults$nums

if(initCCA) datB <- CCAresults$biomass_ages
if(initNEUS) datB <- NEUSresults$biomass_ages
if(initNOBA) datB <- NOBAresults$biomass_ages


# now we need a weight at age to convert numbers to weight
# if I do biomass_age$atoutput divided by nums$atoutput I should have wt@age, yes?

nums_test <- with(datN, aggregate(atoutput, list(species, agecl), sum))
names(nums_test) <- c("species", "agecl", "nums")

bio_test  <- with(datB, aggregate(atoutput, list(species, agecl), sum))
names(bio_test) <- c("species", "agecl", "wt")

calcwtage <- merge(bio_test, nums_test) %>%
  mutate(wtAtAge=wt/nums)

wtagecheck <- ggplot(calcwtage, aes(x=agecl, y=wtAtAge)) +
  geom_point() + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=1, scales="free")

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=4, scales="free")


```

```{r surveyNbased, eval=FALSE}

# this uses result$nums so will need a weight at age to get biomass

if(initCCA) datN <- CCAresults$nums
if(initNEUS) datN <- NEUSresults$nums
if(initNOBA) datN <- NOBAresults$nums

survey_testNall <- create_survey(dat = datN,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# as above, make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

# variable name needs to be wtAtAge or doesnt work
wtage <- calcwtage %>%
  select(species, agecl, wtAtAge)

surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
#save for later use, takes a long time to generate
saveRDS(surveyB_fromN, file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))

if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
if(initNEUS) NEUSsurveyB_fromN <- surveyB_fromN
if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN

```

```{r readsurvey2}

if(initCCA) {
  CCAsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
}

if(initNEUS) {
  NEUSsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
}

if(initNOBA) {
  NOBAsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
}

```

Now we plot the results of the numbers-based (census) survey biomass against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above. 

```{r matchBfromN, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; first 9 species." }

if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
if(initNEUS) surveyB_fromN <- NEUSsurveyB_fromN
if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN

#all species comparison, time intervals hardcoded for NEUS and NOBA
# WHOOPS! Fixed for CCA but generalize this!
compareB_N <-ggplot() +
  geom_line(data=surveyB_fromN, aes(x=time,y=atoutput, color="survey census N->B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 

```

```{r matchBfromN2, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 

```

```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 

```

```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 

```

```{r matchBfromN5, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species. "}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free") 

```

We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion. It may be best to have the survey index for testing derived directly from the true biomass output. 

